---
title: "What is the impact of utility companies' charging station rebates on EV adoption in Maryland?"
format: gfm
editor: visual
---

Step 1. Install necessary packages.

```{r}
install.packages("tidyverse")
install.packages("kableExtra")
```

Step 2. Declare that you will use these packages in this session.

```{r}
library("tidyverse")
library("kableExtra")
```

Step 3. Upload the dataframe that you have created in Spring 2024 into the repository.

Step 4. Open the dataframe into the RStudio Environment.

```{r}
df<-read.csv("panel (1).csv")
```

Step 5. Use the **head** and **kable** function showcase the first 10 rows of the dataframe to the reader.

```{r}
kable(head(df))
```

## Question 1: What is the frequency of this data frame?

Answer: Semi-annually from 2016 - 2023

## Question 2: What is the cross-sectional (geographical) unit of this data frame?

Answer: Zipcode

Step 6. Use the **names** function to display all the variables (column) in the dataframe.

```{r}
names(df)

```

## Question 3: Which column represents the treatment variable of interest?

Answer: Rebate columns for all electric companies

## Question 4: Which column represents the outcome variable of interest?

Answer: Vehicle.Count (EV Sales)

Step 7: Create a boxplot to visualize the distribution of the outcome variable under treatment and no treatment.

```{r}
df2 <- df %>%
  mutate(Registration.Date = as.Date(Registration.Date)) %>%
  filter(Registration.Date < as.Date("2022-06-01")) %>%
  group_by(Registration.Date, NAME, AEC_Rebate, PSEG_Rebate, AEC, PSEG) %>%
  summarize(Total_EV = sum(Vehicle.Count))

ggplot(df2, aes(x = Registration.Date, y = Total_EV, color = NAME)) + geom_line()

model1<-lm(Total_EV ~ AEC*AEC_Rebate + PSEG*PSEG_Rebate, data=df2)
summary(model1)
```

Step 8: Fit a regression model $y=\beta_0 + \beta_1 x + \epsilon$ where $y$ is the outcome variable and $x$ is the treatment variable. Use the **summary** function to display the results.

```{r}

glimpse(df)

#df2 <- df %>%
 # mutate(Registration.Date = as.Date(Registration.Date)) %>%
  #filter(Registration.Date < as.Date("2022-06-01")) 

```

## Question 5: What is the predicted value of the outcome variable when treatment=0?

Answer: 36 EV sales per zip code per 6 months

## Question 6: What is predicted value of the outcome variable when treatment=1?

Answer: AEC = 28(decreased), PSEG = 103(increased)

## Question 7: What is the equation that describes the linear regression above? Please include an explanation of the variables and subscripts.

Answer: $$
y = \beta_0 + \beta_1 AEC + \beta_2 AECRebate + \beta_3 AEC*AECRebate + \beta_4  + \beta_5 PSEG + \beta_6 PSEGRebate + \beta_7 PSEG*PSEGRebate
$$

y represents the expected amount of vechile sales in the zip codes covered by AEC and PSEG in New Jersey

## Question 8: What fixed effects can be included in the regression? What does each fixed effects control for? Please include a new equation that incorporates the fixed effects.

Answer: Fixed effects include Zipcode, Company, Semi-annual Indicator, Year

\## Question 9: What is the impact of the treatment effect once fixed effects are included?

```{r}
#install.packages("lfe")
library("lfe")


df4 <- df2 %>%
  filter(NAME == "Jersey Central Power & Light" | NAME == "Public Service Electric & Gas Co.")
  
model3 <- felm(Total_EV ~ PSEG*PSEG_Rebate, data=df4)
summary(model3)

df5 <- df2 %>%
  filter (NAME == "Atlantic City Electric" | NAME == "Rockland Electric Company")

model4 <- felm(Total_EV ~ AEC*AEC_Rebate, data=df5)
summary(model4)

```

Answer:

# Questions for Week 5

## Question 10: In a difference-in-differences (DiD) model, what is the treatment GROUP?

Answer: Zip codes covered by AEC and PSEG

## Question 11: In a DiD model, what are the control groups?

Answer: Zip codes covered by OR and JCPL

## Question 14: What are the next steps of your research?

Step 9: Change the document format to gfm

Step 10: Save this document as README.qmd

Step 11: Render the document. README.md file should be created after this process.

Step 12: Push the document back to GitHub and observe your beautiful document in your repository!

Step 13: If your team has a complete dataframe that includes both the treated and outcome variable, you are done with the assignment. If not, make a research plan in Notion to collect data on the outcome and treatment variable and combine it into one dataframe.
