---
title: "What is the impact of utility companies' charging station rebates on EV adoption in New Jersey?"
format: gfm
editor: visual
---

## Introduction:

Electric vehicle (EV) adoption is starting to rise and one of the most challenging aspects is the transition between gasoline-powered vehicles to electric ones. In the state of New Jersey, many utility companies are starting to offer rebates and incentives for the purpose of installing EV charging stations in various counties to maintain widespread accessibility (Qmerit, 2024). The rebates are for reducing financial issues and burdens on businesses, property wonders, and other organizations in order to make it easier to install these stations. Lowering the costs and increasing the availability of charging stations through rebate programs, this attracts more EV customers and supports the current ones. Our research is important to society because New Jersey contains a high urban environment (New Jersey’s cleanenergy program, 2020). These charging station rebates lead to clean transportation choices and offer awareness to the public about the importance of positive EV consumption. Using this research to promote property wonders and businesses to understand the relationship between rebates and EV adoption, helps New Jersey’s objectives to reducing greenhouse gasses and promoting healthy energy uses.

## Existing Research:

Electric vehicle (EV) charging rebates are financial incentives that are given by governments and certain organizations to encourage people to acquire electric vehicles. When one applies for a rebate after purchasing an EV charging station, one can receive some money back. This helps promote cleaner transportation options and makes charging stations more affordable and convenient.This study will build on existing findings and delve deeper into the impact of utility companies' charging station rebates on electric vehicle adoption. Despite there being some research done on similar topics, there are noticeable gaps. Certain studies have focused on border areas from a geographical standpoint, or also lack specific information on the impact of rebates on EV adoption rates in New Jersey. Other studies also may not have researched consumer behavior, accessibility affordability, and rebate mechanisms overall. Muehlegger & Rapson (2022) studied EV adoption in California and helped strengthen our data when focusing on New Jersey by using reliable comparisons (Muehlegger & Rapson, 2022). The state of California gives a broader perspective of what the subsidies and purchases of EVs are from another geographic location. Clinton & Steinberg (2019) explained the relationship between financial incentives and EV adoption through its lessons (Clinton & Steinberg, 2019). Our study will address those gaps due to focusing strictly on New Jersey and its direct impact of the rebates on EV adoption.

## Data Description:

### Outcome Variable:

-   The outcome variable is Vehicle.Count (EV sales) in the zip codes covered by AEC and PSEG in New Jersey. This measures the number of electric vehicles registered, which serves as a proxy for EV adoption in response to the treatment (rebate programs).

### Treatment Variable:

-   The treatment variable is the rebate programs offered by the utility companies (AEC_Rebate, PSEG_Rebate). These rebates are intended to incentivize the installation of EV charging stations, which, in turn, could increase EV adoption.

-   **PSEG:**  Electric Vehicle Charging Program which was implemented on June 15, 2022, offers \$1,500 for Level 2 EVSE make-ready infrastructure. \$5,000 discount for utility-side make-ready costs.

-   **Atlantic City Electric:** Residential Charger Rebate Program implemented on February 17, 2021, rebate up to 50% of electrical upgrade costs – or a maximum of \$1,000 – for installing a Level 2 EV charger at home.

### Frequency/Geographic Unit:

-   Frequency: The dataset is semi-annual, with observations collected every six months from 2016 to 2023.

-   Geographic Unit: The data is organized by zip code.

### Treated and Untreated Group:

-   Treated Group: Zip codes served by utility companies that provide rebates, specifically AEC and PSEG.

-   Untreated Group: Zip codes served by utility companies that do not provide rebates, specifically OR (Rockland Electric Company) and JCP&L (Jersey Central Power & Light).

### Data Cleaning and Analysis:

Step 1. Install necessary packages.

```{r}
install.packages("tidyverse")
install.packages("kableExtra")
```

Step 2. Declare that you will use these packages in this session.

```{r}
library("tidyverse")
library("kableExtra")
```

Step 3. Upload the dataframe that you have created in Spring 2024 into the repository.

Step 4. Open the dataframe into the RStudio Environment.

```{r}
df<-read.csv("panel (1).csv")
```

Step 5. Use the **head** and **kable** function showcase the first 10 rows of the dataframe to the reader.

```{r}
kable(head(df))
```

Step 6: Create a boxplot to visualize the distribution of the outcome variable under treatment and no treatment.

```{r}
df2 <- df %>%
  mutate(Registration.Date = as.Date(Registration.Date)) %>%
  filter(Registration.Date < as.Date("2022-06-01")) %>%
  group_by(Registration.Date, NAME, AEC_Rebate, PSEG_Rebate, AEC, PSEG) %>%
  summarize(Total_EV = sum(Vehicle.Count))

JCPLandPSEG <- df2 %>%
  filter(NAME == "Public Service Electric & Gas Co." |
         NAME == "Jersey Central Power & Light")

ACEandREC <- df2 %>%
  filter(NAME == "Atlantic City Electric" | 
         NAME == "Rockland Electric Company")

ggplot(df2, aes(x = Registration.Date, y = Total_EV, color = NAME)) + geom_line()

ggplot(JCPLandPSEG, aes(x = Registration.Date, y = Total_EV, color = NAME)) + geom_line() + geom_vline(xintercept = as.Date("2021-06-15"), linetype = "dashed") + theme(legend.position = c(0.3, 0.8)) + 
  scale_color_manual(values = c("Public Service Electric & Gas Co." = "#F77708", 
  "Jersey Central Power & Light" = "#54b6CC")) +
  labs(y="Total EV Sales", x="Date", color = "Utility Companies")

ggplot(ACEandREC, aes(x = Registration.Date, y = Total_EV, color = NAME)) + geom_line() + geom_vline(xintercept = as.Date("2021-02-01"), linetype = "dashed") + theme(legend.position = c(0.3, 0.8)) + scale_color_manual(values = c("Atlantic City Electric" = "#54b6CC", 
  "Rockland Electric Company" = "#F77708")) +
  labs(y="Total EV Sales", x="Date", color = "Utility Companies")
 
model1<-lm(Total_EV ~ AEC*AEC_Rebate + PSEG*PSEG_Rebate, data=df2)
summary(model1)
```

## Preliminary Regression Results:

Step 8: Fit a regression model $y=\beta_0 + \beta_1 x + \epsilon$ where $y$ is the outcome variable and $x$ is the treatment variable. Use the **summary** function to display the results.

```{r}

glimpse(df)

#df2 <- df %>%
 # mutate(Registration.Date = as.Date(Registration.Date)) %>%
  #filter(Registration.Date < as.Date("2022-06-01")) 

```

## Question 1: What is the predicted value of the outcome variable when treatment=0?

Answer: 1146 EV sales per electric company region per 6 months

## Question 2: What is predicted value of the outcome variable when treatment=1?

Answer: AEC: 1146 + 576 + 2709 + -145 = 4286

PSEG: 1146 + 13321 - 343 + 17469 = 31593

## Question 3: What is the equation that describes the linear regression above? Please include an explanation of the variables and subscripts.

Answer: $$
y = \beta_0 + \beta_1 AEC + \beta_2 AECRebate + \beta_3 AEC*AECRebate + \beta_4  + \beta_5 PSEG + \beta_6 PSEGRebate + \beta_7 PSEG*PSEGRebate
$$

y represents the expected amount of vechile sales in the zip codes covered by AEC and PSEG in New Jersey

```{r}
#install.packages("lfe")
library("lfe")


df4 <- df2 %>%
  filter(NAME == "Jersey Central Power & Light" | NAME == "Public Service Electric & Gas Co.")
  
model3 <- felm(Total_EV ~ PSEG*PSEG_Rebate, data=df4)
summary(model3)

df5 <- df2 %>%
  filter (NAME == "Atlantic City Electric" | NAME == "Rockland Electric Company")

model4 <- felm(Total_EV ~ AEC*AEC_Rebate, data=df5)
summary(model4)

```

## Conclusion:

In conclusion, we examined four electricity companies, Public Service Electric & Gas (PSEG), Atlantic City Electric (ACE), Rockland Electric Company (REC), and Jersey City Electric and Power (JCPL), in the state of New Jersey to see how their rebate programs implemented the adoption of electric vehicles. To be more specific, we compared ACE to REC and JCPL to PSEG, as the pairs had similar trends in the number of EVs in zipcodes covered by them.

Starting with ACE and REC, the AEC rebate was implemented in 2021 before REC's in 2022, so we decided to compare the increase in EVs caused by the AEC rebate. After running the model, we found our intercept to be 1078, which would've been the amount of EV sales in REC's region per 6 months without any rebates, and the number of EVs in ACE's region to be 1078 + 644 = 1722. Now, with the AEC rebate, the model predicted the number of EV sales to in AEC's region to grow to 1078 + 644 + 1209 + 1182 = 4181, while the number of EV's in REC's region to grow to 1078 + 1209 = 2287.

Continuing with PSEG and JCPL, the PSEG rebate was also implemented in 2021 before JCPL's in 2022, so we decided to compare the increase in EVs caused by the PSEG rebate. After running the model, we found our intercept to be 8819, which would've been the amount of EV sales in JCPL's region per 6 months without any rebates, and the number of EVs in PSEG's region to be 8819 + 5950 = 14,769 before its rebate was implemented. Then, once the PSEG rebate was implemented, the model predicted the number of EV sales in PSEG's region to grow to 8819 + 5950 + 12009 + 7524 = 34302 EVs while the number of EV's in JCPL's region to grow to 8819 + 12009 = 20828 EVs.

## Future Plans:

Our future plans include a more detailed analysis taking other variables into account. For example, we could look at electricity costs in different seasons to see if that impacts the perceived benefit of EV charging station rebates. We also plan to look at other states and rebate policies based on the data available. This would help us compare our preliminary results of New Jersey to other states to see if trends are similar.

## Bibliography:

Clinton, B. C., & Steinberg, D. C. (2019). Providing the Spark: Impact of financial incentives on battery electric vehicle adoption. Journal of Environmental Economics and Management, 98, 102255. https://doi.org/10.1016/j.jeem.2019.102255 

Electric Vehicle Incentive Programs \| NJ OCE Web Site. (2020). Njcleanenergy.com. https://www.njcleanenergy.com/ev?gad_source=1&gclid=Cj0KCQiA_qG5BhDTARIsAA0UHSIhzHJDWQ_QdSyRdiJZgAbI6xi1JgecGyTVxcU6rW_Zgq9IrJDr1wAaAh9FEALw_wcB 

EV Residential Charging Program. EV Residential Charging Program - PSE&G. (n.d.). https://nj.myaccount.pseg.com/myservicepublic/electricvehicles-residential-program

EV promotions & tax credits. (n.d.).https://jerseycentral.chooseev.com/promos/#:\~:text=Through%20JCP&L's%20EV%20Driven%20Program,at%20each%20qualifying%20installation%20site

Muehlegger, E., & Rapson, D. S. (2022). Subsidizing low- and middle-income adoption of electric vehicles: Quasi-experimental evidence from California. Journal of Public Economics, 216(216), 104752. https://doi.org/10.1016/j.jpubeco.2022.104752 

New Jersey EV Rebates & Charging Incentives (2024, July). Qmerit. https://qmerit.com/location/new-jersey/ 

Orange & Rockland. (n.d.). https://www.oru.com/en/our-energy-future/electric-vehicles/new-jersey/residential-ev-drivers/charger-ready-for-home

Residential Charger Rebate. Atlantic City Electric. (n.d.). https://www.atlanticcityelectric.com/smart-energy/innovation-technology/residential-charger-rebate
